---
title: "Bus_problem"
author: '"Akash","Sohan","Anvit","Sujana"'
date: "2023-09-04"
output:
  pdf_document: default
  html_document: default
---

#### Using the following model for a bus-ridership analysis

#### At each stop, each passenger is likely to get off the bus independent of others with a 20% chance

#### At every stop, there is a 50%/40%/10% chance of 0/1/2 passengers getting on board

#### Bus is empty when it arrives at the first stop


```{r}
entering_bus = function() {
  
  return(sample(c(0,1,2),size = 1,prob = c(0.50,0.40,0.10)))
}

exiting_bus = function(number_prev){
  
  return(sum(sample(c(1,0),size =number_prev ,replace = TRUE, prob = c(0.20,0.80))))
}

```


#### Code to calculate using simulation, the probability that the bus is empty after visiting the 10th stop


```{r}
bus_10Passenger = function() {
  bus_seat=c()
  for (j in c(1:10)) {
    if (j == 1) {
      bus_seat=c(bus_seat,entering_bus())
    }
    else if(bus_seat[j-1]==0){
    bus_seat=c(bus_seat,entering_bus()) 
    }
    else {
    bus_seat=c(bus_seat,bus_seat[j-1]+entering_bus()-exiting_bus(bus_seat[j-1]))
    }
  }
  return(bus_seat)
}

```


#### Simulating 10000  people entering and exiting the bus

```{r}
nsimuationss=10000
simulatedData = replicate(nsimuationss,bus_10Passenger())
```


#### Count of 0 entries at 10th stop

```{r}
table(simulatedData[10,])
```


#### Bus is empty at 10th Stop


```{r}
sum(simulatedData[10,]==0)/length(simulatedData[10,])
```

