---
title: "disranvar"
output: html_document
date: "2023-10-03"
---

#An oil companany has 20% chance of striking oil when drilling a well. What is the probability the company strikes 5 oil out of next 10 attempts?

```{r}
n=10
p=0.2
dbinom(5,n,p)
```
# An oil companany has 20% chance of striking oil when drilling a well. What is the probability the company strikes 5 oil out of next 10 attempts?

```{r}
n=10
p=0.2
sum(dbinom(0:5,n,p))
pbinom(5,n,p)

```
# Approximately 42% of people have type o blood. On a given day in a blood bank, 120 people arrive to donate blood. What is the probability that 30 of those 120 ppl have type o blood?

```{r}
n=120
p=0.42
j=30 #P(x=30)  #Refer notes
dbinom(j,n,p)

Nsuccess = 42000
Nfailure = 58000
n=120
j=30
dhyper(j,Nsuccess,Nfailure,n)

```
# An oil companany has 20% chance of striking oil when drilling a well. What is the probability the company drills 7 times to strike oil the 3rd time?

```{r}
dbinom(3,4,0.99)
dhyper(3,9900,100,4) #cummulative prob #without replacement

```

```{r}
#(1) Sample size is small compared to population size (condition 1 not satisfied)
#(2) Sample 
dbinom(4,9000,0.4)
dhyper(4,4000,6000,9000)

```

```{r}
p = 0.2
r = 3
j = 7
dbinom(j-r,j,p) #point prob #j-r if binomial is in negative


```

```{r}
p = 0.2
r = 20
j = 100
dbinom(r,j,p)

```
# 48/50 Defective items in a box. 
```{r}

nsuccess = 48
nfailure = 2
n = 45
1-dhyper(0,nsuccess,nfailure,n)


n = 45
p = 48/50
1-dbinom(0,n,p)
```
# Telephone Survey
```{r}

p = 0.1
r = 4
j = 10
dbinom(j-r,j,p)
```
#stoplight
```{r}
p = 0.31
n = 8
j=5:n
#dbinom(j,n,p)+dbinom(j+1,n,p)+dbinom(j+2,n,p)+dbinom(j+3,n,p) or
sum(dbinom(j,n,p))
```

```{r}
j=4
pbinom(j,n,p) # x<=4, P(x-0)+....P(x-4)
1-pbinom(j,n,p) #cummulative distributive function
```

```{r}
n=4
p=4/10
j=2
dbinom(j,n,p)   # Exactly 2
pbinom(j,n,p)   # At most 2  # j=0,1,2
j=1             # x<=1
1-pbinom(j,n,p) # At least 2

```

```{r}
simulatedData = rbinom(1000,n,p)
simulatedData
mean(simulatedData == 2) # Exactly
dbinom(2,n,p)
mean(simulatedData <= 2) # At Most
pbinom(2,n,p)
mean(simulatedData >= 2) # At Least
1-pbinom(1,n,p)


```
```{r}
library(ggplot2)
library(dplyr)
```


```{r}
n=4
p=0.4
nsimulations=1e5
```


```{r}
simulations=rbinom(nsimulations,n,p)
table(simulations)

```
```{r}
dfBinom = as.data.frame(table(simulations))
dfBinom

```
```{r}
colnames(dfBinom) = c('j','Freq')
dfBinom['Prob'] = dfBinom['Freq'] / nsimulations

```

```{r}
p = ggplot(data = dfBinom) + 
    geom_col(aes(x = j, y = Prob), width =0.1, fill = 'blue') +
    ggtitle('Probability Mass Function (PMF) of X~Bin(n = 4, p = 0.4)') +
    theme_minimal()
p

```

```{r}

dfBinom['Cumprob'] = cumsum(dfBinom['Prob'])
dfBinom['Cumprob']

```
```{r}
dfBinom
```


```{r}
colnames(dfBinom) = c('j','Freq')
dfBinom['Prob'] = dfBinom['Freq'] / nsimulations
dfBinom["CumProb"]=cumsum(dfBinom['Prob'])

```

```{r}
q = ggplot(data = dfBinom) + 
    geom_col(aes(x = j, y = CumProb), width =0.1, fill = 'red') +
    ggtitle('Cummulative Distributive Frequency (CDF) of X~Bin(n = 4, p = 0.4)') +
    theme_minimal()
q

```

```{r}
n=4
p=0.4
nsimulations=1e5
simulatedData=rbinom(nsimulations,n,p)
mean(simulatedData)
sum(simulatedData) / nsimulations
```


```{r}
(0*(sum(simulatedData == 0)/nsimulations) + 1*(sum(simulatedData == 1)/nsimulations) + 
2*(sum(simulatedData == 2)/nsimulations) + 3*(sum(simulatedData == 3)/nsimulations) +
4*(sum(simulatedData == 4)/nsimulations))

```


```{r}
n=40
j=3
Nsuccess=4
Nfailure=48
dhyper(j,Nsuccess,Nfailure,n)

```
```{r}


```

