---
title: "Discrete Random Variables"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(dplyr)
```

---

**Scenario**: A salesman has scheduled two appointments to sell software, one in the morning and another one in the afternoon. There are two software editions available: the base edition costing Rs. 5000 and the premium edition costing Rs. 10000. His morning appointments typically lead to a sale with a 30% chance while the afternoon ones typically lead to a sale with a 60% chance independent of what happened in the morning. If the morning appointment ends up in sale, the salesman has a 70% chance of selling the premium edition and if the afternoon appointment ends up in a sale, he is equally likely to sell either of the editions. Let X be the random variable representing the total Rupee value of sales. What are the different values that X can take? Calculate the probability that X takes the value 5000?

---


```{r}
# Sampling space for appointment success (0 corresponds to no sale, 1 corresponds to a sale)
s_appointment = c(0,1)

# Appointment success and failure probabilities
p_morning = 0.3 # Success probability of sales in morning
p_afternoon =  0.6 # Success probability of sales in afternoon
p_appointment = matrix(nrow = 2, ncol = 2, c(1-p_morning, p_morning, 1-p_afternoon, p_afternoon), byrow = TRUE)

# Sampling space for software type
s_software = c(0,1)

# Software type probabilities
p_software = matrix(nrow = 2, ncol = 2, c(0.3,0.7,0.5,0.5), byrow = TRUE)

# Function that simulates one trial of the random experiment which is
# what the salesman earns on a random day
salesResult = function(){
  result = numeric(2)
  # Simulate whether sales happen in morning and afternoon appointments
  for (j in c(1:2)){
    result[j] = sample(s_appointment, size = 1, replace = TRUE, prob = p_appointment[j, ])
  }
  earnings = ifelse(result[1] == 1, sample(s_software, size = 1, replace = TRUE, prob = p_software[1, ]), 0) + ifelse(result[2] == 1, sample(s_software,size=1, replace = TRUE, prob = p_software[2, ]), 0)
  return(earnings)
}
```


```{r}
# Number of simulations
nsimulations = 1e5
simulatedData = replicate(nsimulations, salesResult())
simulatedData
```


```{r}
# Probability that the salesman earns Rs 5000
mean(simulatedData == 5000)

# Can we calculate all the probabilities in one shot?
table(simulatedData) / nsimulations

```

```{r}
as.data.frame(simulatedData)
```

```{r}
dfsales = as.data.frame(simulatedData)
head(dfsales,5)
```

```{r}
dfsales = as.data.frame(simulatedData)
colnames(dfsales)=c('Earnings')
head(dfsales)
```

```{r}
p = ggplot(data=dfsales) + geom_bar(aes(x=Earnings))
p
```

```{r}

```

```{r}
p = ggplot(data = dfsales) + geom_bar(aes(x = Earnings), fill = 'red')
p
```

```{r}
dfsales = as.data.frame(table(simulatedData))
head(dfsales)
```

```{r}
dfsales = as.data.frame(table(simulatedData))
head(dfsales)
colnames(dfsales)=c('Earnings','Frequency')
head(dfsales)
```

```{r}
dfsales['Probability'] = dfsales['Frequency'] / nsimulations
head(dfsales)
```

```{r}
dfsales = as.data.frame(table(simulatedData))
head(dfsales)
colnames(dfsales)=c('Earnings','Frequency')
head(dfsales)
dfsales['Probability'] = dfsales['Frequency'] / nsimulations
head(dfsales)

```


```{r}
p = ggplot(data = dfsales) + geom_col(aes(x = Earnings,Probability), fill = 'blue')
p
```

```{r}
p = ggplot(data = dfsales) + geom_col(aes(x = Earnings,Probability), fill = 'green', width = 0.5)
p
```




