---
title: "Graphing with ggplot"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(ggplot2)
```

#### Read Data

```{r}
foodData = read.csv('food-texture.csv',
                    header = TRUE, row.names = 1)
head(foodData)
```

```{r}
str(foodData)
```

```{r}
nrow(foodData)
ncol(foodData)
rownames(foodData)
colnames(foodData)         
```

#### Accessing elements of the dataframe

```{r}
foodData$Oil
foodData[['Oil']]
foodData[, 'Oil']
foodData['Oil']
```

#### Factor Data structure in R

```{r}
gender = c('male','female','female','male','female','male','male')
str(gender)
print(gender)
gender_categorical=factor(gender)
print(gender_categorical)
str(gender_categorical)
levels(gender_categorical)
gender_categorical[1]

```

#### Add a new column called taste with levels (1) poor (2) fair (3) good (4) excellent

```{r}
#foodData['taste']=sample(c('poor','fair','good','excellent'), nrow(foodData), replace = TRUE, p = c(0.25,0.25,0.25,0.25))
#head(foodData)
```

#### write the updated dataframe to food-texture.csv

```{r}
#write.csv(foodData,'food-texture.csv')

```

#### create list of continuous and categorial feature names

```{r}
continuous_cols = c('Oil','Density','Fracture','Hardness')
categorical_cols = c('Crispy', 'taste')
```

#### convert all categorical coulmns (that we know) into factor type in one shot

```{r}
foodData[categorical_cols]
```


```{r}
foodData[categorical_cols] = lapply(foodData[categorical_cols],factor)
str(foodData)
```

#### create a ggplot canvas

```{r}
P = ggplot(data = foodData)
P

```

#### visualize and understand the relationship between two continous variables using color coding (scatter plot): Oil and Density

```{r}
P = ggplot(data = foodData) + geom_point(aes(x = Density, y = Oil, color = Crispy))
P

```

```{r}
P = ggplot(data = foodData) + geom_point(aes(x = Density, y = Oil, size = Crispy))
P

```

#### visualize and understand the relationship between a continous and  a categorial feature(box plot): Density and Crispy

```{r}
P = ggplot(data = foodData) + geom_boxplot(aes(x = Crispy, y = Density))
P
```

#### Understand what is inisde a box plot

```{r}
#foodData$Density
#foodData$Crispy==10
#foodData[foodData$Crispy == 10, ]
#foodData[foodData$Crispy == 10 & foodData$Oil <= 17,]
density10 = foodData[foodData$Crispy == 10, 'Density']
density11 = foodData[foodData$Crispy == 11, 'Density']
#density11 = foodData[foodData$Crispy == 10, c('Density','Hardness')]
density10
density10

```

```{r}
density10                                     #vector
mean(density10)
density10 - mean(density10)                   # mean centered value
(density10 - mean(density10))^2               # squared mean centered
mean((density10 - mean(density10))^2)         # variance of the density
sqrt(mean((density10 - mean(density10))^2))   # Standard Deviation
```

```{r}
var(density10)
sd(density10)
```

```{r}
var(dfDensity10$Density)
sd(dfDensity10$Density)
```

```{r}
dfDensity10 = data.frame(density10)
colnames(dfDensity10) = c('Density')
dfDensity10$DensityMC = density10 - mean(density10)
dfDensity10$DensityMCs = (dfDensity10$DensityMC)^2
str(dfDensity10)
```

```{r}
#Component plot of mean centered density values of samples with crispy level 10
p = ggplot(data = dfDensity10) +
    geom_point(aes(x = c(1:nrow(dfDensity10)), y = DensityMC), color = 'blue') +
    #geom_line(aes(x = c(1:nrow(dfDensity10)), y = Density), color = 'green') +
    geom_line(aes(x = c(1:nrow(dfDensity10)), y = mean(DensityMC)), color = 'red') +
    labs(x = 'Sample Number', y = 'Density (kg/m^3)', title = "Component plot of Density")
  
p
```

```{r}
#
lapply(foodData[continuous_cols], scale)
```



